---
title: "PhyloTrace pyMLST Workflow"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE, warning = FALSE)
# Derive project root from the Rmd file location, works both when knitting and interactively
rmd_dir <- if (interactive()) {
  dirname(rstudioapi::getSourceEditorContext()$path)
} else {
  dirname(knitr::current_input(dir = TRUE))
}
root <- normalizePath(file.path(rmd_dir, ".."))
knitr::opts_knit$set(root.dir = root)
source(file.path(root, "bin/pymslt.R"))

db_path <- file.path(root, "pseudomonas_aeruginosa.db")
genome_path <- "/home/marianfreisleben/Desktop/ncbi_dataset/subset2"
```

## 1. Download cgMLST Scheme

Download the *Pseudomonas aeruginosa* cgMLST scheme from PubMLST via `wgMLST import`. The scheme is seeded into the SQLite database at `db_path`.

```{r download-scheme, eval=FALSE}
download_cgmlst_scheme(
  scheme = "Pseudomonas aeruginosa",
  db_path = db_path
)
```

## 2. Inspect the Database Structure

Read all tables from the SQLite database into a named list and inspect the structure. The two key tables are `sequences` (allele sequences) and `mlst` (per-isolate allele assignments).

```{r read-database}
database <- read_database(db_path)
str(database)
```

## 3. Type Isolates

Run BLAT-based cgMLST typing against the database for one or more genome FASTA files. `type_genomes` calls the `loop-pymlst.sh` wrapper and parses its output for QC flags (duplicate, species mismatch, success).

```{r type-genomes, eval=FALSE}
database <- type_genomes(
  database = database,
  db_path = db_path,
  genome_input = genome_path
)
```

## 4. Add SHA-256 Sequence Hashes

Hash every allele sequence in `database$sequences` using SHA-256 via `openssl`. The resulting `sha256` column provides scheme-independent allele identifiers that are stable across database versions.

```{r add-hashes}
database <- add_sequence_hashes(database)
str(database$sequences)
```

## 5. Retrieve Allelic Profiles

Pivot the MLST table into a wide format where each row is one isolate (`souche`) and each column is a locus, with SHA-256 hashes as values.

```{r allele-profiles}
profiles <- get_isolate_allele_profiles(database)
head(profiles[, 1:10])
```

## 6. Synchronize with the Remote Database

Write all in-memory changes (including the new `sha256` column) back to the SQLite file on disk.

```{r synchronize, eval=FALSE}
synchronize_database(database, db_path)

# Check if hashes were saved
database <- read_database(db_path)
names(database$sequences)
```

